Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-19 paper/paper(1).qmd
Add files via upload

The Canadian electoral process, epitomizing democratic principles, empowers citizens to elect parliamentarians who shape Parliament's composition and, consequently, the government formation.
In the recent 2021 federal election, Justin Trudeau's Liberal Party secured a majority in the 44th Parliament[@Nakhaie_2006].
Voters' primary concerns include the cost of living, healthcare, climate change, post-pandemic recovery, and immigration, with dissatisfaction towards Trudeau's government prevalent.
Trudeau's popularity lags behind UK's Poilievre[@Nakhaie_2006].
The government's ambitious immigration plan aims to admit 500,000 migrants annually by 2025, albeit facing a backlog.
The Conservative Party, with a focus on economic prosperity, emphasizes attracting highly skilled immigrants while addressing concerns like living costs and housing issues[@Elections_2023].

@@

To enhance control and prevention of illegal immigration, the Liberal Party has suggested modifications to the Immigration and Refugee Protection Act.
They propose establishing an annual refugee quota of 250, specifically aimed at supporting human rights defenders, journalists, and humanitarian workers seeking safety and resettlement in Canada.
Looking ahead, the Liberal Party intends to incrementally increase this refugee quota, aiming for refugees to comprise 15% of all immigrants.
Conversely, the share of economic immigrants is planned to decrease to 58%, while the family reunification immigrant ratio will remain constant.
These proportions are expected to be maintained over the next two years.
Moreover, the Canadian national government recently unveiled an ambitious plan to admit 500,000 migrants annually by 2025, looking for nearly 1.5 million newcomers in Canada within the next three years (Canada's Immigration Levels Plan 2024-2026).
Historically, the Liberal government's immigration strategy tends to be more permissive, offering broader entry routes compared to that of the Conservative Party[@Nakhaie_2006].
Based on our analysis of historical trends, governmental policies, and electoral procedures, our hypothesis aligns with the prevailing expectation that the Conservatives will emerge as the likely winners in the 2025 election.
**We anticipate that the Liberals will likely secure the highest number of votes, with the Conservatives expected to closely follow in second place.** Our research indicates that the vote-counting process in Canada significantly influences the outcome, as voters do not directly elect a specific prime minister candidate.
Initially, all votes are tallied based on first preferences.
If a candidate obtains 50% of the vote, they win.
If not, the candidate with the fewest votes is eliminated[@Elections_2023].
Then, votes from supporters of the eliminated candidate are redistributed to their second-choice candidate.
This iterative redistribution continues until one candidate achieves 50% of the vote.
Hence, our objective is to forecast the election probabilities for the Conservative and Liberal parties and compare which is higher[@Elections_2023]. The estimand of our research is relationship between age, gender, province, religion, income and education and probability of log odds of vote Conservative or Liberal party through binary logistic regression model.

@@

In this paper, I used binary logistic regression models to forecast the likelihood of individuals affiliating with the Conservative or Liberal parties.Binary logistic regression is a statistical technique tailored for predicting binary outcomes, where the variable under consideration has two possible results.
In our study, these outcomes represent the probability of individuals aligning with each of the two parties.Each party's likelihood will be predicted through separate logistic regression models, with the response variables indicating whether individuals vote for the corresponding party or not.
The predictors used in these models include age, gender, province, education, religion, and income before tax.
This regression approach is particularly advantageous when there's a nonlinear relationship between independent and dependent variables, especially in scenarios with categorical dependent variables like ours.
Additionally, we will incorporate post-stratification in our analysis to enhance accuracy.
Post-stratification helps refine precision estimates by ensuring that the sample is more representative of the entire population.
We utilize this technique to predict the probability of voter shares for the Liberal or Conservative parties in the upcoming election.

# Data

@@

# Model

@@

# Result

@@

# Discussion

>>>>

The Canadian electoral process, epitomizing democratic principles, empowers citizens to elect parliamentarians who shape Parliament's composition and, consequently, the government formation.
In the recent 2021 federal election, Justin Trudeau's Liberal Party secured a majority in the 44th Parliament[@Nakhaie_2006].
To enhance control and prevention of illegal immigration, the Liberal Party has suggested modifications to the Immigration and Refugee Protection Act.
They propose establishing an annual refugee quota of 250, specifically aimed at supporting human rights defenders, journalists, and humanitarian workers seeking safety and resettlement in Canada.
Looking ahead, the Liberal Party intends to incrementally increase this refugee quota, aiming for refugees to comprise 15% of all immigrants.
Conversely, the share of economic immigrants is planned to decrease to 58%, while the family reunification immigrant ratio will remain constant.
These proportions are expected to be maintained over the next two years.
Moreover, the Canadian national government recently unveiled an ambitious plan to admit 500,000 migrants annually by 2025, looking for nearly 1.5 million newcomers in Canada within the next three years (Canada's Immigration Levels Plan 2024-2026).
Historically, the Liberal government's immigration strategy tends to be more permissive, offering broader entry routes compared to that of the Conservative Party[@Nakhaie_2006]

@@

Based on our analysis of historical trends, governmental policies, and electoral procedures, our hypothesis aligns with the prevailing expectation that the Conservatives will emerge as the likely winners in the 2025 election.
**We anticipate that the Liberals will likely secure the highest number of votes, with the Conservatives expected to closely follow in second place.** Our research indicates that the vote-counting process in Canada significantly influences the outcome, as voters do not directly elect a specific prime minister candidate.
Initially, all votes are tallied based on first preferences.
If a candidate obtains 50% of the vote, they win.
If not, the candidate with the fewest votes is eliminated[@Elections_2023].
Then, votes from supporters of the eliminated candidate are redistributed to their second-choice candidate.
This iterative redistribution continues until one candidate achieves 50% of the vote.
Hence, our objective is to forecast the election probabilities for the Conservative and Liberal parties and compare which is higher[@Elections_2023]. The estimand of our research is relationship between age, gender, province, religion, income and education and probability of log odds of vote Conservative or Liberal party through binary logistic regression model.

@@

In this paper, in @sec-Model, I used binary logistic regression models to forecast the likelihood of individuals affiliating with the Conservative or Liberal parties.Binary logistic regression is a statistical technique tailored for predicting binary outcomes, where the variable under consideration has two possible results.
In @sec-Model, these outcomes represent the probability of individuals aligning with each of the two parties.Each party's likelihood will be predicted through separate logistic regression models, with the response variables indicating whether individuals vote for the corresponding party or not.
The predictors used in these models include age, gender, province, education, religion, and income before tax is discussed in @sec-Data.
This regression approach is particularly advantageous when there's a nonlinear relationship between independent and dependent variables, especially in scenarios with categorical dependent variables like ours, details can be found in @sec-Data.
Additionally, in @sec-Model, we will incorporate post-stratification in our analysis to enhance accuracy.
Post-stratification helps refine precision estimates by ensuring that the sample is more representative of the entire population.
We utilize this technique to predict the probability of voter shares for the Liberal or Conservative parties in the upcoming election in @sec-Result.
Finally, the consideration of potential bias and limitations is discussed in @sec-Discussion.

# Data{#sec-Data}

@@

# Model{#sec-Model}

@@

# Result{#sec-Result}

@@

# Discussion{#sec-Discussion}
